---
title: "03_results.Rmd"
author: "Adam Hanbury-Brown"
date: "3/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model 1; Shive + 2018 (-April 1 snowpack)
```{r pressure, echo=FALSE}
mod1_predictors <- c("recoveryTime",'slope',"aspect","CWDhist","AEThist","postFireSAP")
#getCorrelationMatrix(mod1_predictors)
mod1 <- fitGAM(df4 %>% filter(postFirePlanting < 0.1,
                              mediumReburns < 0.01,
                              lowReburns < 0.01),
               mod1_predictors,response = "RRI_end", scaleData = T)
plot(mod1)
summary(mod1)
print(paste("AIC",AIC(mod1)))
```
Results of Mod 1: 
1. Everything is significant expect aspect. So it pretty much works.
2. Generally the relationships are pretty reasonable (except CWD). Could remove CWD because its correlated with AET and doesn't give a meaningful relationship.
3. Not clear why SAP is negative at low SAP
4.

AIC is lower than if all terms are linear
AIC is slightly higher if you use adjusted folded northness instead of aspect 
If I use conCov here instead of RRI end then SAP_400 has a more reasonable result

##Model 2. Stewart et al
```{r,echo=F}
mod2_predictors <- c("recoveryTime","slope","aspect","pptYr1_3_annual_mean","burnSev")
#getCorrelationMatrix(mod1_predictors)
mod2 <- fitGAM(df4 %>% filter(postFirePlanting < 0.1,
                              mediumReburns < 0.01,
                              lowReburns < 0.01),
               mod2_predictors,response = "RRI_end", scaleData = T)
plot(mod2)
summary(mod2)
print(paste("AIC",AIC(mod2)))
```




